% ============================================================
%  §3  Method
% ============================================================
\label{sec:method}

\subsection{Problem Formulation}
\label{sec:method_formulation}

Let $\mathcal{D}_{\text{train}} = \{(x_i, y_i, \mathbf{v}_i)\}_{i=1}^{k}$ denote a small training
subset of $k$ cases, where $x_i$ is the question, $y_i$ the ground-truth answer, and $\mathbf{v}_i$
the associated visual input (one or more images).
Let $\pi_\theta$ be a frozen VLM backbone (no gradient updates permitted).
We define a capability set $\mathcal{C} = \{\mathcal{S}, \mathcal{T}\}$, where
$\mathcal{S}$ is a library of \emph{skills} (structured reasoning SOPs) and
$\mathcal{T}$ is a library of \emph{tools} (executable Python programs).
An agent $f_{\mathcal{C}}$ operates by retrieving relevant capabilities from $\mathcal{C}$,
applying them during problem solving, and invoking $\pi_\theta$ for reasoning.
Our goal is to learn $\mathcal{C}$ from $\mathcal{D}_{\text{train}}$ to maximise performance on a
held-out validation set $\mathcal{D}_{\text{val}}$, subject to the constraint
$\nabla_\theta = 0$ (no weight updates).

\subsection{Evolution Loop}
\label{sec:method_loop}

Figure~\ref{fig:hero} (left) illustrates the \sysname{} evolution loop, which processes each
training case as follows.

\textbf{(1) Attempt.}
Agent $f_{\mathcal{C}}$ attempts case $(x_i, \mathbf{v}_i)$ using the current capability set
$\mathcal{C}$.  Relevant skills are retrieved by semantic similarity to $x_i$; relevant tools are
identified by their applicability SOPs (see Section~\ref{sec:method_mastery}).

\textbf{(2) Evaluate.}
The attempted answer is compared to $y_i$.  If correct, the loop advances to the next case.
If incorrect, a failure record
$\delta_i = (x_i, y_i, \mathbf{v}_i, \hat{y}_i, \tau_i)$ is created, where
$\hat{y}_i$ is the wrong answer and $\tau_i$ is the agent's reasoning trace.

\textbf{(3) Diagnose.}
The AnalyzerDecider role receives the failure record $\delta_i$ together with the original images
$\mathbf{v}_i$ and any intermediate tool outputs from the failed attempt.
It produces a structured root-cause analysis and a decision
$a \in \{\texttt{gen\_skill},\, \texttt{gen\_tool},\, \texttt{gen\_both},\, \texttt{give\_up}\}$.
Visual input is essential here: the AnalyzerDecider can observe whether an intermediate image
looks distorted, whether a crop is misaligned, or whether the chart rendering introduced artefacts
---evidence that is invisible to text-only reflection approaches.

\textbf{(4) Generate.}
The Generator role produces the requested capability.
Skill generation outputs a Markdown SOP; tool generation outputs Python code
(Section~\ref{sec:method_skill} and~\ref{sec:method_tool}).

\textbf{(5) Validate and Promote.}
Generated capabilities are validated before promotion.
A skill is validated by re-attempting $\delta_i$ with the new SOP included.
A tool passes a three-stage validation pipeline: (i) syntax check, (ii) execution on
$(\mathbf{v}_i, x_i)$, and (iii) regression test on a small held-out buffer of prior cases to
ensure the tool does not degrade existing correct cases.
Capabilities that pass validation are added to $\mathcal{C}$; failed capabilities are discarded.

\textbf{(6) Retry.}
After promoting new capabilities, the agent retries $\delta_i$ with the updated $\mathcal{C}$.
This retry serves as an additional quality signal but does not affect $\mathcal{D}_{\text{val}}$
evaluation, which uses only the \emph{frozen} capability set produced after all iterations.

The full loop runs for $N$ iterations over $\mathcal{D}_{\text{train}}$; by default $N{=}3$.
Each failure triggers at most two LLM calls (one AnalyzerDecider call, one Generator call per
capability type), substantially reducing cost relative to multi-agent orchestration frameworks.

\begin{algorithm}[t]
\SetAlgoLined
\KwIn{Training set $\mathcal{D}_{\text{train}}$, frozen VLM $\pi_\theta$,
      iterations $N$, training size $k$}
\KwOut{Capability set $\mathcal{C} = (\mathcal{S}, \mathcal{T})$}
$\mathcal{C} \leftarrow (\emptyset, \emptyset)$\;
\For{$n = 1$ \KwTo $N$}{
  \ForEach{$(x_i, y_i, \mathbf{v}_i) \in \mathcal{D}_{\text{train}}$}{
    $\hat{y}_i, \tau_i \leftarrow f_{\mathcal{C}}(x_i, \mathbf{v}_i;\, \pi_\theta)$\;
    \If{$\hat{y}_i \neq y_i$}{
      $\delta_i \leftarrow (x_i, y_i, \mathbf{v}_i, \hat{y}_i, \tau_i)$\;
      $a,\, r \leftarrow \textsc{AnalyzerDecider}(\delta_i, \mathbf{v}_i;\, \pi_\theta)$\;
      \If{$a \in \{\texttt{gen\_skill}, \texttt{gen\_both}\}$}{
        $s \leftarrow \textsc{Generator}(r;\, \pi_\theta)$\;
        \lIf{\textsc{Validate}$(s, \mathcal{D}_{\text{train}})$}{$\mathcal{S} \leftarrow \mathcal{S} \cup \{s\}$}
      }
      \If{$a \in \{\texttt{gen\_tool}, \texttt{gen\_both}\}$}{
        $t \leftarrow \textsc{Generator}(r;\, \pi_\theta)$\;
        \If{\textsc{Validate}$(t, \mathbf{v}_i, \mathcal{D}_{\text{train}})$}{
          $t.\text{mastery} \leftarrow \textsc{Mastery}(t, \mathcal{D}_{\text{train}};\, \pi_\theta)$\;
          $\mathcal{T} \leftarrow \mathcal{T} \cup \{t\}$\;
        }
      }
    }
  }
}
\Return{$\mathcal{C}$}\;
\caption{\sysname{} Evolution Loop}
\label{alg:main}
\end{algorithm}

\subsection{Visual Failure Diagnosis}
\label{sec:method_diagnosis}

The AnalyzerDecider is the core reasoning component that distinguishes \sysname{} from text-only
self-improvement methods.
It receives as input: (a) the original images $\mathbf{v}_i$, (b) any intermediate tool outputs
produced during the failed attempt (cropped sub-images, processed renders), (c) the agent's full
reasoning trace $\tau_i$, and (d) the current capability set $\mathcal{C}$ (to avoid regenerating
existing capabilities).

From these inputs, the AnalyzerDecider produces a structured diagnosis comprising:
\begin{itemize}[noitemsep]
  \item \textbf{Root cause}: a description of why the agent failed, grounded in visual evidence
        (e.g.\ ``the bar chart labels are too small to read reliably without OCR preprocessing'');
  \item \textbf{Bottleneck type}: cognitive (the agent had the right percepts but applied the wrong
        strategy) vs.\ perceptual (the agent could not extract the necessary visual information);
  \item \textbf{Next action} $a$: one of \texttt{gen\_skill}, \texttt{gen\_tool},
        \texttt{gen\_both}, or \texttt{give\_up} (when the failure is attributed to inherent
        model limitations with no recoverable capability).
\end{itemize}

The visual channel is critical for accurate diagnosis.
In our experiments, cases where the AnalyzerDecider had access to images produced
diagnoses that correctly identified perceptual bottlenecks \textcolor{red}{[TBD]\%} of the time,
vs.\ \textcolor{red}{[TBD]\%} for text-only diagnosis (Section~\ref{sec:ablation}).

\subsection{Skill Evolution}
\label{sec:method_skill}

A \emph{skill} in \sysname{} is a structured Markdown document with four mandatory fields:
\textbf{When-to-Use} (a trigger condition expressed as a natural-language predicate over the
question and image type),
\textbf{Strategy} (a numbered step-by-step SOP for approaching the problem class),
\textbf{Common Pitfalls} (failure modes to avoid), and
\textbf{Worked Example} (a solved instance illustrating the SOP).

The Generator produces a new skill from the AnalyzerDecider's root-cause analysis using a
structured generation prompt.
If a skill covering the same problem class already exists in $\mathcal{S}$, the Generator
instead \emph{merges}: it appends new insights (additional pitfalls, alternative strategy
branches) to the existing skill, preventing library bloat while preserving accumulated knowledge.
Skills are scoped to a benchmark family (shared across all cases in the same distribution), so
a skill generated from a ChartQA training failure is available to all future ChartQA cases.

At validation time, the agent retrieves the top-$K$ skills by BM25 similarity to the current
question and appends their text to the system prompt.
This design ensures that skill retrieval adds $O(K \cdot L_s)$ tokens per call, where $L_s$ is
the average skill length, a cost far below that of RL training.

\subsection{Tool Evolution and Progressive Mastery}
\label{sec:method_tool}
\label{sec:method_mastery}

A \emph{tool} in \sysname{} is a Python function (at most 150 lines) that takes one or more
image paths as input and returns processed images or extracted data.
Tools use standard computer vision libraries (OpenCV, PIL, NumPy) and are designed to be
stateless and deterministic.

\textbf{Tool generation.}
Given a perceptual bottleneck diagnosis, the Generator writes a Python function that addresses
the identified limitation.
Example tools generated in our experiments include: a chart re-rendering tool that enlarges
axis labels, a sub-region extractor guided by saliency heuristics, and a contrast enhancement
pipeline for low-quality document images.

\textbf{Three-stage validation.}
Before promotion, each tool passes:
\begin{enumerate}[noitemsep, label=(\roman*)]
  \item \emph{Syntax check}: the code compiles and executes without errors;
  \item \emph{Origin test}: the tool improves the agent's answer on the failure case $\delta_i$
        that triggered its creation;
  \item \emph{Regression test}: the tool does not reduce accuracy on a buffer of 20 prior cases
        from $\mathcal{D}_{\text{train}}$ that the agent currently handles correctly.
\end{enumerate}
This staged validation rejects tools that are too specialised (pass (i) and (ii) but fail (iii))
or trivially wrong (fail (i)).

\textbf{Progressive mastery.}
After a tool passes three-stage validation, \sysname{} runs a mastery phase.
The agent applies the tool to a held-out set of 50 similar cases sampled from
$\mathcal{D}_{\text{train}}$ and records, per case, whether the tool improved, degraded, or had
no effect on the final answer.
From this profiling run, the agent synthesises a \emph{mastery SOP}: a Markdown document
describing the tool's supported input patterns (image types where it helps),
negative patterns (image types where it hurts or produces no gain), and recommended trigger
conditions.
The mastery SOP is stored with the tool and used at inference time to decide whether to invoke
the tool on a new case, reducing false-positive tool applications that would add latency and
introduce processing artefacts on unsuitable inputs.

This mastery phase is what distinguishes \sysname{} from prior tool-creation methods~\citep{tang2023creator,cai2024latm}: rather than generating tools and deploying them indiscriminately, the agent learns the boundary conditions of each tool's effectiveness, enabling selective, context-appropriate invocation.

\subsection{Tool Mastery from External Tool Sets (\sysname{} Mode B)}
\label{sec:method_mastery_ext}

The evolution loop above (Mode A) both \emph{generates} tools and \emph{profiles} their
applicability.  In settings where a practitioner already supplies a useful tool set
$\mathcal{T}_0 \neq \emptyset$, \sysname{} can operate in \textbf{Mode B}: skip tool generation
entirely and apply the mastery phase directly to the provided tools.

\textbf{Formal statement.}
Given an external tool set $\mathcal{T}_0 = \{t_1, \ldots, t_m\}$ and training data
$\mathcal{D}_{\text{train}}$, run the mastery profiling procedure (Section~\ref{sec:method_mastery})
for each $t_j \in \mathcal{T}_0$ to produce mastery SOPs
$\{t_j.\text{mastery}\}_{j=1}^{m}$.
Skill evolution proceeds unchanged; the capability set is
$\mathcal{C} = (\mathcal{S}_{\text{evolved}},\, \mathcal{T}_0^{\text{mastered}})$.

Mode B requires no code generation and fewer LLM calls (one mastery profiling pass per tool),
making it practical when tool quality is already high but deployment strategy is unknown.
Experiment II (Section~\ref{sec:exp2}) evaluates this mode.
