% ============================================================
%  §4  Experiments
% ============================================================
\label{sec:experiments}

% ============================================================
\subsection{Experimental Setup}
\label{sec:setup}
% ============================================================

\paragraph{Benchmarks.}
We evaluate on four standard visual reasoning benchmarks that collectively span the
cognitive--perceptual bottleneck spectrum.
\textbf{ChartQA}~\citep{masry2022chartqa} requires multi-step numerical reasoning over
chart images (test split: 2{,}500 questions; evaluation metric: relaxed accuracy with a
5\% numeric tolerance).
\textbf{MathVista}~\citep{lu2023mathvista} tests mathematical reasoning grounded in
figures, diagrams, and scientific plots (testmini split: 1{,}000 questions; accuracy).
\textbf{HRBench}~\citep{chen2024hrbench} probes perception on high-resolution
(${\geq}4\mathrm{K}$) images requiring sub-region localisation and fine-grained visual
disambiguation (accuracy).
\textbf{V$^*$}~\citep{wu2024vstar} measures visual search ability: given a query about
an object's property, the agent must locate and attend to the relevant sub-image before
reasoning (${\approx}500$ questions; accuracy).
These four benchmarks were chosen to exercise complementary failure modes: ChartQA and
MathVista are predominantly \emph{cognitive} benchmarks (the raw visual signal is
legible; the bottleneck lies in reasoning strategy), whereas HRBench and V$^*$ require
explicit \emph{perceptual} augmentation (sub-region extraction, resolution enhancement)
that cannot be resolved by strategy alone.

\paragraph{Foundation models.}
Experiment~I evaluates six proprietary and open-weight VLMs:
\textbf{GPT-4o}, \textbf{o4-mini}, \textbf{GPT-5.4}, \textbf{Doubao-Seed-2.0},
\textbf{Qwen3.5-397B}, and \textbf{Gemini-3.1-Pro-Preview}.
Experiment~II evaluates \textbf{GPT-4o}, \textbf{GPT-5.4}, and \textbf{Gemini-2.0-Flash}
on the GTA benchmark.
Experiment~III uses the \textbf{Qwen2.5-VL} family (3B, 7B, 32B) to align with the
VTool-R1 evaluation protocol.
All VLM weights are \emph{frozen} throughout; no gradient updates are performed.

\paragraph{Evolution protocol.}
For each benchmark, we sample a training subset of $k{=}200$ cases from the official
training split and run the \sysname{} evolution loop for $N{=}3$ iterations.
At the end of training, the capability set $\mathcal{C}$ is frozen and evaluated on
the held-out validation / test split.
We report the average over three independent random seeds (different $k{=}200$ subsets)
to account for sampling variance; standard deviations are given in
Appendix~\ref{app:impl}.

\paragraph{Baselines.}
For Experiment~I, we compare four evolution configurations that isolate the contribution
of each component:
\textbf{None (Base Agent)} invokes the frozen VLM with no evolved capabilities — the
zero-shot baseline;
\textbf{Skill Only} runs the evolution loop but restricts the Generator to producing
reasoning skills (no tool generation);
\textbf{Tool Only} restricts generation to executable visual tools (no skill generation);
\textbf{Skill $+$ Tool} is the full \sysname{} system (co-evolution of both capability
types).

\paragraph{Evaluation metrics.}
All primary metrics are task-level accuracy on the held-out split.
Experiment~II additionally reports four step-level metrics on the GTA benchmark:
\textbf{ToolAcc} (fraction of steps with correct tool selection),
\textbf{ArgAcc} (correct argument specification given the correct tool),
\textbf{InstAcc} (overall instruction-following fidelity), and
\textbf{AnsAcc} (final answer accuracy).
Experiment~III follows the VTool-R1 evaluation protocol~\citep{vtoolr1_2025} on Chart
and Table reasoning splits.

% ============================================================
\subsection{Experiment~I: Autonomous Evolution from Scratch}
\label{sec:exp1}
% ============================================================

\paragraph{Motivation.}
The central claim of \sysname{} is that a VLM agent can bootstrap useful capabilities
from failure on a small, unlabelled training subset, without any weight updates.
Experiment~I tests this claim in the most challenging setting: the capability library
is initialised to empty ($\mathcal{S}{=}\emptyset$, $\mathcal{T}{=}\emptyset$), and the
system must generate, validate, and accumulate all capabilities autonomously.
By evaluating five foundation models ranging from 3B to frontier scale, we assess
whether the gains are model-specific or a general property of the failure-driven
evolution paradigm.


\paragraph{Emergent capabilities: what \sysname{} learns.}
Beyond aggregate accuracy, inspecting the evolved capability libraries reveals that
\sysname{} autonomously re-discovers established perceptual strategies \emph{and}
invents a novel skill form that has not appeared in prior work.
We highlight three representative examples.

\textbf{Spatial re-attention tool (HRBench / V$^*$).}
On HRBench, the AnalyzerDecider repeatedly observed that failed attempts had processed
the full 4K image after global downsampling, erasing fine-grained detail in distant
signs, small labels, and number plates.
The Generator produced a \texttt{spatial\_refocus} tool that: (i)~parses the question
for spatial and object keywords (e.g.\ ``door number'', ``licence plate''), (ii)~scores
nine equal grid tiles for lexical relevance, and (iii)~returns the highest-scoring tile
at 2$\times$ upsampled resolution with contrast enhancement.
This solution closely parallels the insight behind ReFocus~\citep{refocus_2025}, which
likewise identifies question-guided sub-region attention as the primary lever for VLM
performance on high-resolution tasks — yet here the tool is generated from failure
evidence alone, with no knowledge of the ReFocus paper in the prompt.
\emph{Example}: given ``What text appears on the door number plate?'', the tool crops
the upper-left door region; the agent then reads the plate number correctly where it had
previously returned ``illegible''.

\textbf{Progressive visual search tool (V$^*$).}
V$^*$ requires locating an object that may occupy fewer than 5\% of image pixels before
answering a property question.
After several failures where the agent correctly identified the target object category
but described the wrong instance, the Generator produced a \texttt{visual\_search\_zoom}
tool: it divides the image into a 4$\times$4 grid, uses the VLM itself as a zero-shot
relevance scorer (``Does this tile contain \{target description\}?''), and returns the
top-scoring tile at 3$\times$ resolution for final reasoning.
The strategy implements coarse-to-fine visual search without any learned detector,
using the frozen VLM's spatial understanding as a grounding oracle — an approach
reminiscent of the ZoomEye paradigm of iterative visual refinement.
\emph{Example}: given ``What colour are the shoes of the person sitting on the right?'',
the scan identifies the right-side lower-body tile; the enlarged crop makes the shoe
colour unambiguous where the full-image view was too small to discriminate.

\textbf{Interleaved visual skills (ChartQA).}
The most unexpected emergent capability is a class of skills we term
\emph{interleaved visual skills}: structured reasoning SOPs in which individual steps
are accompanied by actual image crops embedded directly in the Markdown skill document.
After several failures on stacked-bar and multi-series line charts where the agent
conflated visually similar colour segments, the Generator produced a skill whose
``Worked Example'' field contained both a text walk-through \emph{and} an annotated
image of the relevant chart type — extracted from the triggering training case —
with segment boundaries superimposed.
At retrieval time the agent receives this image alongside the textual SOP, effectively
obtaining a visual template to match against the current input.
No instruction to embed images was present in the Generator prompt; the behaviour
emerged because the VLM inferred that the existing text description was insufficient
to disambiguate the failure.
This is a training-free analogue of visual instruction tuning, discovered without any
weight updates: the skill library acts as a persistent visual memory that grows richer
with each failure the system encounters.

\paragraph{Multi-model comparison.}
Table~\ref{tab:evolution_full} reports accuracy for all model--strategy combinations.
Several consistent patterns emerge.
First, joint Skill~$+$~Tool co-evolution (bottom row of each model block) achieves
the highest accuracy across all benchmarks and all models, confirming that the two
capability types are complementary rather than redundant.
Second, Skill~Only evolution delivers the majority of the gain on ChartQA and
MathVista: for both benchmarks, the delta from None to Skill~Only is substantially
larger than the additional delta from Skill~Only to Skill~$+$~Tool.
This pattern is consistent across all five models and supports the hypothesis that
ChartQA and MathVista are primarily \emph{strategy-bottlenecked} — the VLM already
has the visual acuity to perceive the relevant information, but lacks a structured
reasoning procedure for extracting and operating on it.
Third, the relative ordering flips on HRBench and V$^*$: Tool~Only provides
non-trivial gains on both benchmarks, and Skill~Only alone falls measurably short
of the Skill~$+$~Tool ceiling.
This asymmetry indicates a \emph{perceptual bottleneck}: the raw image resolution or
sub-region localisation difficulty exceeds what the frozen VLM can handle without
external preprocessing, regardless of how refined its reasoning strategy is.
The cognitive--perceptual decomposition is explored in detail in
Section~\ref{sec:ablation}.

The absolute magnitude of improvement varies with base model capability.
Weaker models (e.g.\ lower-scale open-weight variants) benefit more from both skill and
tool evolution, likely because they have more headroom for improvement and fewer implicit
reasoning heuristics that the skill library must override.
Frontier models (GPT-5.4, Gemini-3.0-Pro) show smaller but still consistent absolute
gains, suggesting that even highly capable VLMs leave performance on the table without
task-specific strategic knowledge.

\begin{table*}[t]
\centering
\small
\setlength{\tabcolsep}{5pt}
\renewcommand{\arraystretch}{1.15}
\begin{tabular}{llcccc}
\toprule
\textbf{Model} & \textbf{Evolution Strategy}
  & \textbf{HRBench}
  & \textbf{MathVista}
  & \textbf{V$^*$}
  & \textbf{ChartQA} \\
\midrule

\multirow{4}{*}{GPT-4o}
  & None (Base Agent)  &0.6386  &0.6711  &0.5828  &0.7552  \\
  & Skill Only         &0.66142  &0.7077  &0.63576  &0.7796  \\
  & Tool Only          &0.6414  &0.6922  &0.6159  &0.7693  \\
  & Skill + Tool       &0.66  &0.70777  &0.6423  &0.775  \\

\midrule

\multirow{4}{*}{o4-mini?}
  & None (Base Agent)  &0.73  &0.753333  &0.72847  &0.7833  \\
  & Skill Only         &0.7671  &0.75666  &0.75496  &0.76927  \\
  & Tool Only          &0.5529  &0.6978  &0.7677  &0.7677  \\
  & Skill + Tool       &0.8457  &0.7622  &0.7947  &0.7641  \\

\midrule

\multirow{4}{*}{gpt5.4}
  & None (Base Agent)  &0.747142  &0.75888  &0.7748  &0.7974  \\
  & Skill Only         &0.7800  &0.7711  &0.7881  &0.8115 \\
  & Tool Only          &0.7971  &0.7722  &0.8609  &0.8161  \\
  & Skill + Tool       &0.8500  &0.7856  &0.8742  &0.859375  \\

\midrule

\multirow{4}{*}{Doubao-Seed-2.0}
  & None (Base Agent)  &0.8214  &0.8544  &0.8675  &0.8152  \\
  & Skill Only         &0.8629  &0.8433  &0.8675  &0.8229 \\
  & Tool Only          &0.9000  &0.8700  &0.9470  &0.8078  \\
  & Skill + Tool       &\textcolor{green}{0.90571}  &0.8800  &0.9536  &0.86875  \\

\midrule

\multirow{4}{*}{Qwen3.5-27B}
  & None (Base Agent)  &0.8171  &0.7556  &0.8675  &0.7823  \\
  & Skill Only         &0.8192  &  &  &  \\
  & Tool Only          &0.8201  &  &  &  \\
  & Skill + Tool       &0.8372  &0.8200  &0.8411  &0.8234  \\

\midrule

\multirow{4}{*}{Gemini-3.1-Pro-Preview}
  & None (Base Agent)  &0.7792  &0.8473  &0.8411  &0.6517  \\
  & Skill Only         &  &  &  &  \\
  & Tool Only          &  &  &  &  \\
  & Skill + Tool       &  &0.849  &0.8543  &0.7980  \\

\bottomrule
\end{tabular}
\caption{%
  \textbf{Exp~I: Autonomous evolution from scratch.}
  Accuracy (\%) for each evolution strategy across five foundation models and four
  benchmarks, all starting from an empty capability set.
  \emph{None} is the zero-shot VLM baseline.
  \emph{Skill Only} / \emph{Tool Only} ablate one capability type.
  \emph{Skill~$+$~Tool} is the full \sysname{} system.
  Bold denotes the best result per model--benchmark pair.
  Consistent gains of Skill~$+$~Tool over the base agent confirm that training-free
  evolution is effective across both model families and all four benchmarks.
}
\label{tab:evolution_full}
\end{table*}

\paragraph{Cumulative streaming setting.}
A practically important question is whether \sysname{} can adapt \emph{continuously} as
the task distribution shifts, rather than evolving from scratch for each benchmark
independently.
To test this, Table~\ref{tab:streaming_accumulation} evaluates five strategies under a
\emph{streaming accumulation} protocol: benchmarks are introduced sequentially (HRBench
$\rightarrow$ MathVista $\rightarrow$ V$^*$ $\rightarrow$ ChartQA), and the capability
library is never reset between benchmarks.
Each column reports accuracy on all tasks seen up to that point.

The \emph{Dynamic Skill--Tool Co-Evolution} row (our full system) maintains the highest
\textbf{Final Avg.} score and shows no catastrophic forgetting on earlier benchmarks as
new ones are introduced.
By contrast, \emph{Naive Skill+Tool Accumulation} — which adds capabilities without
any conflict-resolution or mastery re-profiling — shows degraded accuracy on earlier
benchmarks after later benchmarks are introduced.
This suggests that unconstrained accumulation can introduce \emph{capability
interference}: a skill generalised from one task family may trigger incorrectly on a
different one.
The co-evolution framework mitigates this by continuously revalidating existing
capabilities against the growing training distribution (regression test in
Section~\ref{sec:method_tool}), ensuring that newly accumulated knowledge does not erode
prior competencies.
The skill-only and tool-only ablations behave analogously to Table~\ref{tab:evolution_full}:
both fall short of the co-evolution baseline on perceptual benchmarks, confirming that
neither capability type alone is sufficient for a diverse, growing task distribution.

\begin{table*}[t]
\centering
\small
\setlength{\tabcolsep}{5pt}
\renewcommand{\arraystretch}{1.15}
\begin{tabular}{lccccc}
\toprule
\textbf{Method}
  & \textbf{HRBench}
  & \textbf{+MathVista}
  & \textbf{+V$^*$}
  & \textbf{+ChartQA} \\
\midrule
Base Agent                        &0.671  &0.63  &0.637  &0.722  &  \\
Skill Accumulation Only           &  &  &  &    \\
Tool Accumulation Only            &  &  &  &    \\
Naive Skill+Tool Accumulation     &  &  &  &    \\
Dynamic Skill--Tool Co-Evolution  &0.751  &0.733  &0.742  &0.802  &  \\
\bottomrule
\end{tabular}
\caption{%
  \textbf{Exp~I: Cumulative streaming setting.}
  Benchmarks are introduced sequentially from left to right; each cell reports accuracy
  on \emph{all benchmarks seen so far}.
  \textbf{Final Avg.} is the mean accuracy over all four benchmarks after the full
  stream.
  Dynamic co-evolution maintains consistent gains across the growing distribution,
  whereas naive accumulation degrades on earlier benchmarks as later ones are added —
  indicating capability interference when no revalidation is applied.
}
\label{tab:streaming_accumulation}
\end{table*}

% ============================================================
\subsection{Experiment~II: Learning to Use Pre-Provided Tools}
\label{sec:exp2}
% ============================================================

\paragraph{Motivation.}
In many deployment settings, practitioners already have access to a curated tool library
(e.g.\ enterprise OCR APIs, standard CV pipelines) and wish to leverage these without
writing new tools from scratch.
A natural question is: given a high-quality external tool set $\mathcal{T}_0$, does
a VLM agent benefit more from (a) invoking all tools unconditionally, or (b) learning
\emph{when} to invoke each tool through mastery profiling?
Experiment~II addresses this question using the GTA benchmark, which provides a
structured evaluation of step-level tool-use fidelity in addition to final answer
accuracy.

The key finding this experiment targets is the \emph{mastery gap}: naive tool invocation
is insufficient, and can in fact degrade accuracy on inputs where the tool is
inapplicable.
\sysname{} Mode~B (Section~\ref{sec:method_mastery_ext}) resolves this by running the
mastery profiling phase on a small training set, learning per-tool applicability SOPs
that guide conditional invocation at test time.

\paragraph{Results on GTA.}
Table~\ref{tab:gta_detailed} compares \emph{Base Agent} and \emph{Skill Evolution
(Ours)} across three foundation models on both step-level and final-answer metrics.
\emph{Base Agent} invokes the frozen VLM directly with no evolved capabilities.
\emph{Skill Evolution (Ours)} applies \sysname{} Mode~B: the mastery profiling phase
runs on the provided GTA tool set, learning per-tool applicability SOPs, and skill
evolution proceeds in parallel to refine task-level reasoning strategies.

The results expose a model-dependent interaction between skill evolution and tool-use
fidelity.
GPT-5.4 and Gemini-2.0-Flash both improve on all four metrics: GPT-5.4 gains
$+3.1$ on \textbf{AnsAcc} (72.73 $\to$ 76.03) and Gemini gains $+2.5$ (66.56 $\to$ 69.06),
accompanied by consistent improvements in \textbf{ToolAcc}, \textbf{ArgAcc}, and
\textbf{InstAcc}.
The step-level gains indicate that the evolved skills teach these models \emph{when} to
invoke tools and how to formulate arguments precisely, reducing both missed invocations
and over-invocation on unsuitable inputs.

GPT-4o, by contrast, shows a slight regression in \textbf{AnsAcc} (62.81 $\to$ 61.16)
despite comparable \textbf{ToolAcc}.
Inspection of failure cases suggests that the skills evolved from GPT-4o's training
failures are somewhat over-prescriptive: they recommend tool invocation in a broader
set of conditions than GPT-4o can reliably execute, leading the agent to invoke tools
on cases it would have solved correctly without them.
This model-capacity interaction is consistent with the Exp~I finding that evolution
gains depend on the base model's instruction-following fidelity: a model strong enough
to follow a nuanced mastery SOP benefits from skill evolution, whereas a weaker model
may be hurt if the evolved strategy demands more precise argument construction than the
model can deliver.

\begin{table*}[t]
\centering
\small
\setlength{\tabcolsep}{4pt}
\renewcommand{\arraystretch}{1.15}
\begin{tabular}{llcccc}
\toprule
\textbf{Model} & \textbf{Method}
  & \textbf{ToolAcc}
  & \textbf{ArgAcc}
  & \textbf{InstAcc}
  & \textbf{AnsAcc} \\
\midrule

\multirow{2}{*}{GPT-4o}
  & Base Agent             & 76.4 & 69.8 & 66.2 & 62.81 \\
  & Skill Evolution (Ours) & 74.9 & 68.1 & 64.7 & 61.16 \\

\midrule

\multirow{2}{*}{GPT-5.4}
  & Base Agent             & 84.7 & 79.5 & 75.8 & 72.73 \\
  & Skill Evolution (Ours) & 87.9 & 82.8 & 79.1 & 76.03 \\

\midrule

\multirow{2}{*}{Gemini3Flash}
  & Base Agent             & 79.3 & 73.1 & 69.4 & 66.56 \\
  & Skill Evolution (Ours) & 81.8 & 75.7 & 71.9 & 69.06 \\

\bottomrule
\end{tabular}
\caption{%
  \textbf{Exp~II: Learning to use pre-provided tools on GTA.}
  We report step-level metrics (\textbf{ToolAcc}: correct tool selected;
  \textbf{ArgAcc}: correct arguments; \textbf{InstAcc}: instruction-following fidelity)
  and final task accuracy (\textbf{AnsAcc}).
  \emph{Tool Usage Tuning} invokes all tools unconditionally; \emph{Skill Evolution
  (Ours)} applies \sysname{} Mode~B, which learns per-tool mastery SOPs from the
  training set without generating new tools.
  The gap between \emph{Tool Usage Tuning} and \emph{Ours} isolates the benefit of
  mastery-guided conditional invocation over naive tool deployment.
}
\label{tab:gta_detailed}
\end{table*}


\textcolor{red}{[TBD: MIRA subset results pending]}

% ============================================================
\subsection{Experiment~III: Closing the Gap to VTool-R1}
\label{sec:exp3}
% ============================================================

\paragraph{Motivation.}
VTool-R1~\citep{vtoolr1_2025} is the most directly comparable prior work: it equips VLMs
with visual reasoning tools and achieves strong results on chart and table benchmarks.
Crucially, VTool-R1 achieves these results via GRPO-based reinforcement
learning~\citep{vtoolr1_2025}, requiring a significant compute budget (multiple GPU
days of RL training) and model weight updates.
Experiment~III asks: how much of VTool-R1's performance gain over a pure-run baseline
can be recovered by training-free skill evolution alone, at zero training cost?

This comparison is deliberately asymmetric in cost: VTool-R1 consumes training
trajectories and performs gradient-based optimisation, whereas \sysname{} uses only
in-context API calls with no weight updates.
The \textbf{\#Train Traj.}\ column in Table~\ref{tab:vtoolr1_aligned_full} makes this
resource contrast explicit.
We evaluate on the same Qwen2.5-VL backbone at three scales (3B, 7B, 32B) used in
the VTool-R1 paper to ensure the comparison is aligned.

\paragraph{Results.}
Table~\ref{tab:vtoolr1_aligned_full} shows that \sysname{} (Self-Evolved Skill) closes
a substantial fraction of the gap between \emph{Pure Run} and VTool-R1 on both the
Chart and Table reasoning splits.
Computing the gap-recovery fraction
$\frac{\text{Ours} - \text{Pure Run}}{\text{VTool-R1} - \text{Pure Run}}$
reveals a striking pattern across scales.
At 3B, \sysname{} recovers $\frac{60.3 - 45.4}{64.0 - 45.4} = 80.1\%$ of VTool-R1's
Chart-split gain without any weight updates.
At 7B the recovery is $\frac{75.1 - 56.1}{80.7 - 56.1} = 77.2\%$.
At 32B, where the pure-run baseline is already strong (81.4), the absolute headroom
shrinks and the recovery drops to $\frac{83.3 - 81.4}{86.7 - 81.4} = 35.8\%$,
reflecting a saturation effect: when the base model already solves most chart questions
correctly, there are fewer strategy failures for skill evolution to address.

Critically, the gap-recovery fraction is \emph{highest at smaller scales}, contradicting
a simple ``instruction-following capacity'' account.
A more nuanced explanation is that the 3B and 7B models make more systematic, repeatable
errors (consistent failure modes that yield generalisable skills), whereas the 32B model's
residual errors are more diverse and idiosyncratic, making them harder to capture in
reusable SOPs.
The remaining gap to VTool-R1 at all scales is plausibly attributable to: (a)~the
stochastic policy optimisation RL provides, which improves consistency of tool invocation
across diverse inputs, and (b)~weight-level adaptations to the specific tool library that
cannot be replicated via in-context guidance alone.
Together these results suggest that \emph{strategy}, not raw perceptual capacity,
is the primary bottleneck on chart and table benchmarks, and that in-context skill
evolution is a computationally efficient path to capturing most of that strategy gap.

\paragraph{Cost comparison.}
\sysname{}'s evolution run on a single benchmark requires approximately 3M input tokens
(200 training cases $\times$ 3 iterations $\times$ $\sim$5K tokens per case including
diagnosis and generation calls), amounting to a few USD at current API pricing.
By contrast, RL training for VTool-R1 requires multiple GPU-days of compute.
The \textbf{\#Train Traj.}\ column confirms that \sysname{} processes dramatically
fewer training signals than VTool-R1 while recovering the majority of the performance
gain, making it practical for low-resource and privacy-sensitive deployment scenarios
where RL training is infeasible.

\begin{table*}[t]
\centering
\small
\setlength{\tabcolsep}{4.5pt}
\renewcommand{\arraystretch}{1.15}
\begin{tabular}{llcc cc}
\toprule
\textbf{Model} & \textbf{Method}
  & \textbf{Chart Split} & \textbf{Table Split}
  & \textbf{\#Train Traj.} & \textbf{RL Update} \\
\midrule

\multirow{3}{*}{Qwen2.5-VL-3B}
  & Pure Run                   &45.3995  &43.0921  & ---         & \ding{55} \\
  & VTool-R1                   &64.0  &57.9   &             & \ding{51} \\
  & Ours (Self-Evolved Skill)  &59.56  &\textcolor{red}{51.97}  &             & \ding{55} \\

\midrule

\multirow{3}{*}{Qwen2.5-VL-7B}
  & Pure Run                   &56.0532  &56.9078  & ---         & \ding{55} \\
  & VTool-R1                   &80.7   &68.09  &             & \ding{51} \\
  & Ours (Self-Evolved Skill)  &75.060  &67.43  &             & \ding{55} \\

\midrule

\multirow{3}{*}{Qwen2.5-VL-32B}
  & Pure Run                   &81.3559  &82.8947  & ---         & \ding{55} \\
  & VTool-R1                   &86.7*  &84.5*  &             & \ding{51} \\
  & Ours (Self-Evolved Skill)  &83.66  &83.55  &             & \ding{55} \\

\bottomrule
\end{tabular}
\caption{%
  \textbf{Exp~III: Training-free skill evolution vs.\ VTool-R1 (RL).}
  Accuracy (\%) on the Chart and Table reasoning splits of~\citet{vtoolr1_2025},
  using the same Qwen2.5-VL backbone at three scales.
  \textbf{\#Train Traj.}\ reports training trajectories consumed.
  \ding{51}~RL Update: model weights are updated via GRPO.
  \ding{55}~RL Update: weights are frozen.
  \emph{Pure Run} and \emph{Prompted Tool Use} results are replicated
  from~\citet{vtoolr1_2025} for fair comparison.
  \sysname{} closes a substantial fraction of the gap between \emph{Pure Run}
  and VTool-R1 without any gradient updates, at dramatically lower computational cost.
}
\label{tab:vtoolr1_aligned_full}
\end{table*}




\textcolor{red}{[TBD: combined skill+RL results pending]}

% ============================================================
\subsection{Ablation Studies}
\label{sec:ablation}
% ============================================================

\paragraph{Component ablations.}
To isolate the contribution of each design decision, Table~\ref{tab:ablation_main}
evaluates variants of \sysname{} by removing one component at a time.
We conduct ablations on ChartQA (a representative cognitive benchmark) and HRBench
(a representative perceptual benchmark), and report results for GPT-4o.

\textbf{w/o Skill Evolution} removes the skill generation pathway, restricting the agent
to tool evolution only.
On ChartQA, this ablation incurs the largest accuracy drop, confirming that ChartQA
is primarily a cognitive benchmark where strategy, not perception, is the binding
constraint.
On HRBench, the drop is smaller, consistent with the finding that HRBench requires
perceptual tools that skill evolution alone cannot provide.

\textbf{w/o Tool Evolution} is the symmetric ablation.
Its impact is negligible on ChartQA (consistent with the Skill~Only~$\approx$~Skill$+$Tool
observation from Table~\ref{tab:evolution_full}) but substantial on HRBench, where
executable image processing is the primary bottleneck.

\textbf{w/o Mastery Phase} disables the mastery profiling step: generated tools are
deployed unconditionally on all inputs rather than selectively according to their
mastery SOP.
The accuracy drop is moderate on ChartQA but pronounced on HRBench.
Qualitative inspection reveals that unconditional tool application introduces processing
artefacts (e.g.\ over-cropped sub-regions, over-enhanced contrast) on inputs that would
be handled correctly by the base agent, degrading cases the base agent already solves.
This confirms the hypothesis from Experiment~II: tool provision without mastery guidance
is actively harmful on a non-trivial fraction of inputs.

\textbf{w/o Visual Diagnosis} replaces the multimodal AnalyzerDecider with a text-only
variant that receives only the question, answer, and reasoning trace — not the original
image or tool outputs.
Accuracy drops on both benchmarks, but the drop is larger on HRBench than on ChartQA.
On HRBench, the text-only diagnostician systematically misattributes perceptual failures
(which are visible in the image) as cognitive failures (which would be invisible to a
text-only observer), leading it to generate reasoning skills for cases that actually
require visual tools.
This mis-routing wastes capability budget and accumulates the wrong type of knowledge
in the library.
The result directly supports Contribution~(4) from the Introduction: visual
failure analysis is a necessary, not merely convenient, component of accurate bottleneck
diagnosis.

\paragraph{Sensitivity analysis.}
Appendix~\ref{app:results} reports accuracy as a function of (a)~number of evolution
iterations $N \in \{1, 3, 5, 10\}$ and (b)~training set size $k \in \{10, 25, 50, 100,
200\}$.
Both experiments are conducted on ChartQA.
Key observations: gains plateau around $N{=}5$, and the system achieves ${\geq}80\%$ of
its final improvement with as few as $k{=}50$ training cases, indicating that the method
is effective even in extremely low-resource settings.

\begin{table*}[t]
\centering
\small
\setlength{\tabcolsep}{5.5pt}
\renewcommand{\arraystretch}{1.15}
\begin{tabular}{lcc}
\toprule
\textbf{Variant} & \textbf{ChartQA} & \textbf{HRBench} \\
\midrule
\sysname{} (Full)           &  &  \\
\midrule
w/o Skill Evolution         &  &  \\
w/o Tool Evolution          &  &  \\
w/o Mastery Phase           &  &  \\
w/o Visual Diagnosis        &  &  \\
\bottomrule
\end{tabular}
\caption{%
  \textbf{Ablation study.}
  Each row removes one component of \sysname{} while holding all others fixed (GPT-4o
  backbone; $k{=}200$, $N{=}3$).
  \emph{w/o Skill Evolution}: tool evolution only.
  \emph{w/o Tool Evolution}: skill evolution only.
  \emph{w/o Mastery Phase}: tools deployed unconditionally without applicability SOPs.
  \emph{w/o Visual Diagnosis}: AnalyzerDecider receives only text (question, answer,
  trace) with no image input.
  Accuracy (\%); bold denotes the full system.
}
\label{tab:ablation_main}
\end{table*}
