% ============================================================
%  §5  Discussion
% ============================================================
\label{sec:discussion}

\paragraph{When does evolution help?}
The bottleneck analysis in Section~\ref{sec:exp1} suggests a practical diagnostic:
skill evolution is effective whenever the benchmark has a recurring family of problem structures
that can be captured in an SOP---which appears to be the common case.
Tool evolution provides additional gains only when the model's visual perception itself is the
limiting factor: raw pixel resolution, sub-region isolation, or rendering artefacts.
A practitioner targeting a new benchmark can use a small pilot run ($k{\approx}25$, $N{=}1$)
to detect which bottleneck type dominates before committing to a full evolution run.

\paragraph{Limitations.}
Several limitations bound the scope of our conclusions.
\textit{(a) Short evolution horizon.}
We report results at $N{=}3$ iterations; the evolution dynamics suggest gains plateau
around $N{=}5$, but we have not evaluated whether longer runs (e.g.\ $N{>}10$) eventually
overfit the training distribution.
\textit{(b) Tool quality ceiling.}
Tool effectiveness is bounded by the VLM's Python coding ability.
For benchmarks requiring specialised computer vision operations beyond standard libraries
(e.g.\ learned feature extractors), the Generator may fail to produce valid tools, limiting
the perceptual gains.
\textit{(c) Benchmark scope.}
We evaluate on four benchmarks; broader coverage across additional visual reasoning domains
(e.g.\ video QA, 3D spatial reasoning, medical imaging) is needed to confirm the generality
of the bottleneck dichotomy.
\textit{(d) API cost.}
Each training case incurs up to $\sim$15K tokens (attempt + failure analysis + generation +
validation), summing to $\sim$3M tokens per benchmark run.  While negligible compared to RL
training, this cost should be accounted for when comparing training-free methods.

\paragraph{Future directions.}
Three extensions merit investigation.
First, \emph{cross-benchmark skill transfer}: skills generated from ChartQA failures may
transfer to MathVista (both involve structured data reading); a shared skill library across
benchmarks could reduce per-benchmark cold-start cost.
Second, \emph{image generation as a reasoning tool}: rather than only processing existing images,
the agent could generate hypothetical visualisations (e.g.\ mentally redrawing a chart in a
clearer format) using image generation APIs, extending the ``think with images'' paradigm
\citep{vtoolr1_2025} to internally imagined percepts.
Third, \emph{online evolution}: rather than a fixed pre-deployment training phase, the agent
could evolve capabilities during deployment, adapting to the distribution shift of live queries.
